### 2017-10-12 writing/paper/sections/coincidences.tex
Such changes

After implementing the First Dominating Pattern Rule we find that there are
\(29\) \(\coinc[(1)]{321}\)-classes, of mesh patterns where the underlying
classical pattern is \(21\). By Note~\NoteRef{note:main} there are
therefore exactly \(29\) \(\coinc{321}\)-classes.

\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern.
Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
\(\coinc[comp]{231}\)-class, but this is not explained by the First Dominating Pattern Rule.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
of elements \(x\) and \(y\). If the region corresponding
to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must therefore contain a (non-empty) decreasing subsequence.
 This gives rise to the following lemma:

\begin{lemma}
    \label{lem:incdec}
    Let \(m =(\sigma, R)\) be a mesh pattern, where the box \((a,b)\) is not
    in \(R\), and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
    \(\pi \preceq m^{\boks{a,b}_a}\)\\(\(\pi \preceq m^{\boks{a,b}_d}\)),
    then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \((a,b)\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
we know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
is an occurrence of \(m_2\). This shows that our two example patterns are coincident.

This result generalises into the following rule for categorising coincidences
of mesh patterns in cases where there is a dominating classical pattern.

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident as a consequence of the Second Dominating Rule if

    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
        \item\label{prop:dom2:cond} Any one of the following four conditions hold \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq \sigma^{\boks{a,b}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R_1\) and \\
                                \((x,b-1)\in R_1 \implies (x,b) \in R_1 \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R_1\) and \\
                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq \sigma^{\boks{a,b}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R_1\) and \\
                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R_1\) and \\
                                  \((x,b-1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
    implies an occurrence of \(m_2\).
    We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
    \(\omega\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).

@@

    Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

This proposition essentially states that we slide all of the points in the box we desire
to shade diagonally, and chose the topmost/bottommost point to replace the original
point in the mesh pattern. \FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\) , the circled points are the same points in the permutation.
\begin{figure}
\begin{tikzpicture}
	\modpattern[3]{1,2}{2,1}{0/0,0/1,0/2,1/0,2/1}
	\filldraw (1.33,1.66) circle (1.5pt);
	\filldraw (1.66,1.33) circle (1.5pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.1,0.9) rectangle (2.5,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.5,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\raisebox{8ex}{\(\coinc{}\)}
\begin{tikzpicture}
	\modpattern[3]{1}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
	\filldraw (2.15,0.85) circle (1.5pt);
	\filldraw (2.3,0.7) circle (1.5pt);
	\draw (2.3,0.7) circle (3pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.4,0.6) rectangle (2.65,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.65,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\caption{A depiction of the second dominating pattern rule using our example patterns.}
\label{fig:d2}
\end{figure}

Together the First Dominating Pattern Rule and the Second Dominating
Pattern Rule fully explain coincidences of classes of the form
\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
mesh patterns, confirming experimental observations.

\subsection{Coincidence classes of Av(\{231, (12, \textit{R})\}).}
When considering the coincidence classes of \(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\)
we first apply the two Dominating Pattern rules previously established.
Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-coincidence classes
reducing to \(59 \coinc[(2)]{231}\)-coincidence classes, considering mesh patterns with
\(12\) as the underlying classical patterns. However we know that there are patterns
where the coincidences are not explained by the rules given above.

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty, and consider choosing the rightmost
point in the region.

This gives us an occurrence of the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

>>>>

After implementing the First Dominating Pattern Rule we find that there are
\(29\) \(\coinc[(1)]{321}\)-equivalence classes, of mesh patterns where the underlying
classical pattern is \(21\). By \NoteRef{note:main} there are
therefore exactly \(29\) \(\coinc{321}\)-equivalence classes.

\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-equivalence
classes between mesh patterns with \(21\) as an underlying classical pattern.
Experimentation shows that there are \(39 \coinc[comp]{231}\)-equivalence
classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
\(\coinc[comp]{231}\)-equivalence class, but this is not explained by the First Dominating Pattern Rule.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
of elements \(x\) and \(y\). If the region corresponding
to the box \(\boks{1,1}\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since the permutation is in \(\av{\perm{2,3,1}}\) this is not possible. This box
must therefore contain a (non-empty) decreasing subsequence.
 This gives rise to the following lemma:

\begin{lemma}
    \label{lem:incdec}
    Let \(m =(\sigma, R)\) be a mesh pattern, where \(\boks{a,b} \notin R\),
    and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
    \(\pi \preceq m^{\boks{a,b}_a}\)\\(\(\pi \preceq m^{\boks{a,b}_d}\)),
    then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \(\boks{a,b}\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
we know that the region corresponding to the box \(\boks{1,1}\) contains a decreasing subsequence.
If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
This example generalises into the following rule.

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), then \(m_1 \coinc{\pi} m_2 \)
    if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom2:cond} Any one of the following four conditions hold \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq \sigma^{\boks{a,b}_a}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \(\boks{a+1,b-1}\notin R_1\) and \\
                                \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1 \) (where \(x\neq a,a+1\)) and\\
                                  \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \(\boks{a-1,b+1}\notin R_1\) and \\
                                  \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
                                  \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq \sigma^{\boks{a,b}_d}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \(\boks{a+1,b+1}\notin R_1\) and \\
                                  \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a,a+1\)) and\\
                                  \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \(\boks{a-1,b-1}\notin R_1\) and \\
                                  \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
                                  \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
    implies an occurrence of \(m_2\).
    We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
    Consider a permutation \(\omega \in \av{\pi}\). Suppose
    \(\omega\) contains \(m_1\) and consider the region corresponding to
    \(\boks{a,b}\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
    of \(m_2\).

@@

    Hence there are no
    points in the region corresponding to the box \(\boks{a,b}\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) can be modified into an occurrence of \(m_2\)
    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

This proposition essentially states that we slide all of the points in the box we desire
to shade diagonally, and chose the topmost/bottommost point to replace the original
point in the mesh pattern. \FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\), the circled points are the same points in the permutation.
\begin{figure}
\begin{tikzpicture}
	\modpattern[3]{1,2}{2,1}{0/0,0/1,0/2,1/0,2/1}
	\filldraw (1.33,1.66) circle (1.5pt);
	\filldraw (1.66,1.33) circle (1.5pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.1,0.9) rectangle (2.5,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.5,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\raisebox{8ex}{\(\coinc{}\)}
\begin{tikzpicture}
	\modpattern[3]{1}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
	\filldraw (2.15,0.85) circle (1.5pt);
	\filldraw (2.3,0.7) circle (1.5pt);
	\draw (2.3,0.7) circle (3pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.4,0.6) rectangle (2.65,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.65,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\caption{A depiction of the second dominating pattern rule using our example patterns.}
\label{fig:d2}
\end{figure}

After implementing the Second Dominating Pattern Rule we find that there are
\(39\) \(\coinc[(2)]{231}\)-equivalence classes, of mesh patterns where the underlying
classical pattern is \(21\). By \NoteRef{note:main} there are
therefore exactly \(39\) \(\coinc{231}\)-equivalence classes.

\subsection{Coincidence classes of Av(\{231, (12, \textit{R})\}).}
When considering the coincidence classes of \(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\)
we first apply the two Dominating Pattern rules previously established.
Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-equivalence classes
reducing to \(59 \coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with
\(12\) as the underlying classical patterns. However there are fewer \(\coinc[comp]{231}\)-equivalence classes.

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are \(\coinc[comp]{231}\)-equivalent.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty, and consider choosing the rightmost
point in the region.
This gives us an occurrence of the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of First Dominating Patter Rule we obtain the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
The highlighted points are an occurrence
of the mesh pattern \(m_2\). This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\omega\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
    is empty then we have an occurrence of \(m_2\).

    By condition~\ref{prop:dom3:condocc} of the proposition there exists
    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}
